---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align = "center",
  fig.retina = 2,
  out.width = "75%",
  dpi = 96
)
knit_hooks$set(pngquant = hook_pngquant)
```


# riskParityPortfolio

[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/riskParityPortfolio)](https://CRAN.R-project.org/package=riskParityPortfolio)
[![CRAN Downloads](https://cranlogs.r-pkg.org/badges/riskParityPortfolio)](https://CRAN.R-project.org/package=riskParityPortfolio)
[![CRAN Downloads Total](https://cranlogs.r-pkg.org/badges/grand-total/riskParityPortfolio?color=brightgreen)](https://CRAN.R-project.org/package=riskParityPortfolio)
[![Rcpp](https://img.shields.io/badge/powered%20by-Rcpp-orange.svg?style=flat)](http://www.rcpp.org/)

[![PyPI version](https://badge.fury.io/py/riskparityportfolio.svg)](https://badge.fury.io/py/riskparityportfolio)
[![Downloads](https://pepy.tech/badge/riskparityportfolio)](https://pepy.tech/project/riskparityportfolio)

[![codecov](https://codecov.io/gh/mirca/riskParityPortfolio/branch/master/graph/badge.svg)](https://codecov.io/gh/mirca/riskParityPortfolio)
[![Travis-CI-Badge](https://travis-ci.org/mirca/riskParityPortfolio.svg?branch=master)](https://travis-ci.org/mirca/riskParityPortfolio)
[![Build status](https://ci.appveyor.com/api/projects/status/dqjti1y461u7sjn8/branch/master?svg=true)](https://ci.appveyor.com/project/mirca/riskparityportfolio/branch/master)
[![CircleCI](https://circleci.com/gh/mirca/riskParityPortfolio.svg?style=svg)](https://circleci.com/gh/mirca/riskParityPortfolio)
[![Docker Build Status](https://img.shields.io/docker/build/mirca/riskparityportfolio.svg)](https://hub.docker.com/r/mirca/riskparityportfolio/)
[![Build Status](https://dev.azure.com/jvmirca/riskParityPortfolio/_apis/build/status/mirca.riskParityPortfolio?branchName=master)](https://dev.azure.com/jvmirca/riskParityPortfolio/_build/latest?definitionId=1&branchName=master)

**riskParityPortfolio** provides tools to design risk parity portfolios.
In its simplest form, we consider the convex formulation with a unique solution proposed by
[Spinu (2013)](https://dx.doi.org/10.2139/ssrn.2297383) and use a cyclical method inspired by
[Griveau-Billion (2013)](https://arxiv.org/pdf/1311.4057.pdf). For more general formulations,
which are usually nonconvex, we implement the successive convex approximation
method proposed by [Feng & Palomar (2015)](https://doi.org/10.1109/TSP.2015.2452219).

The latest stable version of **riskParityPortfolio** is available at https://CRAN.R-project.org/package=riskParityPortfolio.

The latest development version of **riskParityPortfolio** is available at https://github.com/dppalomar/riskParityPortfolio.

**Check out the documentation here: [https://mirca.github.io/riskParityPortfolio](https://mirca.github.io/riskParityPortfolio).**

## Installation
To install the latest stable version of **riskParityPortfolio** from CRAN, run the following commands in R:
```{r, eval = FALSE}
> install.packages("riskParityPortfolio")
```

To install the development version of **riskParityPortfolio** from GitHub, run the following commands in R:
```{r, eval = FALSE}
> install.packages("devtools")
> devtools::install_github("dppalomar/riskParityPortfolio")
```

To get help:
```{r, eval = FALSE}
> library(riskParityPortfolio)
> help(package = "riskParityPortfolio")
> package?riskParityPortfolio
> ?riskParityPortfolio
```

Please cite **riskParityPortfolio** in publications:
```{r, eval = FALSE}
> citation("riskParityPortfolio")
```

You can also get **riskParityPortfolio** from Docker as follows:
```
$ docker pull mirca/riskparityportfolio
```

#### Microsoft Windows
On MS Windows environments, make sure to install the most recent version of
**Rtools**.

### Python

A Python3 implementation of the vanilla method is available in PYPI and can be installed as follows:
```
$ pip install riskparityportfolio
```

Alternatively, the development version can be installed as
```
$ git clone https://github.com/dppalomar/riskParityPortfolio
$ cd python
$ pip install -e .
```

## Usage of **riskParityPortfolio**

### R

```{r}
library(riskParityPortfolio)

set.seed(42)
# create covariance matrix
N <- 5
V <- matrix(rnorm(N^2), ncol = N)
Sigma <- cov(V)

# risk parity portfolio
res <- riskParityPortfolio(Sigma)
names(res)
res$w
res$risk_contribution
c(res$w * (Sigma %*% res$w))

# risk budggeting portfolio
res <- riskParityPortfolio(Sigma, b = c(0.4, 0.4, 0.1, 0.05, 0.05))
res$risk_contribution/sum(res$risk_contribution)
```

### Python
```{python}
import numpy as np
import riskparityportfolio as rpp
np.random.seed(42)

# creates a correlation matrix from time-series of five assets
x = np.random.normal(size=1000).reshape((5, -1))
corr = x @ x.T

# create the desired risk budgeting vector
b = np.ones(len(corr)) / len(corr)

# design the portfolio
w = rpp.design(corr, b)
print(w)

# compute the risk budgeting
rc = w @ (corr * w)
print(rc / np.sum(rc))

# let's try a different budget
b = np.array([0.01, 0.09, .1, .1, .7])
w = rpp.design(corr, b)
print(w)
rc = w @ (corr * w)
print(rc / np.sum(rc))
```


## Documentation
For more detailed information, please check the
[vignette](https://CRAN.R-project.org/package=riskParityPortfolio/vignettes/RiskParityPortfolio.html).

## Citation
If you find this package useful in your research, please consider citing the following works:

- J. V. de M. Cardoso and D. P. Palomar (2019). riskParityPortfolio:
  Design of Risk Parity Portfolios. R package version 0.1.2.
  <https://CRAN.R-project.org/package=riskParityPortfolio>
- Y. Feng, and D. P. Palomar (2015). SCRIP: Successive Convex Optimization Methods for
  Risk Parity Portfolio Design. _IEEE Trans. on Signal Processing_, vol. 63, no. 19,
  pp. 5285-5300. <https://doi.org/10.1109/TSP.2015.2452219>
- F. Spinu (2013). An Algorithm for Computing Risk Parity Weights.
  <https://dx.doi.org/10.2139/ssrn.2297383>
- T. Griveau-Billion, J. Richard, and T. Roncalli (2013). A fast algorithm for computing High-dimensional risk parity portfolios. <https://arxiv.org/pdf/1311.4057.pdf>


## Contributing

We welcome all sorts of contributions. Please feel free to open an issue
to report a bug or discuss a feature request.

## Links
Package: [CRAN](https://CRAN.R-project.org/package=riskParityPortfolio) and [GitHub](https://github.com/dppalomar/riskParityPortfolio).

README file: [CRAN-readme](https://CRAN.R-project.org/package=riskParityPortfolio/readme/README.html)
and [GitHub-readme](https://github.com/dppalomar/riskParityPortfolio/blob/master/README.md).

Vignettes: [CRAN-html-vignette](https://CRAN.R-project.org/package=riskParityPortfolio/vignettes/RiskParityPortfolio.html),
[CRAN-pdf-vignette](https://CRAN.R-project.org/package=riskParityPortfolio/vignettes/RiskParityPortfolio-pdf.pdf), and
[R/Finance 2019 slides](https://docs.google.com/viewer?url=https://github.com/dppalomar/riskParityPortfolio/raw/master/vignettes/RFinance2019-slides.pdf).
