---
title: |
  | Simple example \vspace{0.2cm}
author: |
  | Daniel P. Palomar \vspace{0.2cm}
  | Hong Kong University of Science and Technology (HKUST) \vspace{0.2cm}
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 1
    latex_engine: xelatex
  bookdown::html_document2:
    base_format: prettydoc::html_pretty
    theme: tactile
    highlight: vignette
    fig_caption: yes
    number_sections: no
    toc: yes
    toc_depth: 2
header-includes:
indent: yes
csl: ieee.csl
bibliography: refs.bib
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.retina = 2,
  out.width = "75%",
  dpi = 96
)
knit_hooks$set(pngquant = hook_pngquant)
# Help on bookdown: https://bookdown.org/yihui/bookdown/
# rmarkdown::render("vignettes/example.Rmd", "bookdown::html_document2")
# rmarkdown::render("vignettes/example.Rmd", "bookdown::pdf_document2")
```

-----------
> This note contains an example under consideration by Alexis Nortier.


# Risk-parity portfolio formulation
The risk-parity portfolio formulation is of the form [@FengPal2015riskparity][@FengPal2016monograph]:
$$\begin{array}{ll}
\underset{\mathbf{w}}{\textsf{minimize}} & R(\mathbf{w})\\
\textsf{subject to} & \mathbf{1}^T\mathbf{w}=1\\
 & \mathbf{w}\ge\mathbf{0},
\end{array}$$
where the risk term is of the form (double summation) $R(\mathbf{w}) = \sum_{i,j=1}^{N}(g_{ij}(\mathbf{w}))^{2}$ or simply (single summation) $R(\mathbf{w}) = \sum_{i=1}^{N}(g_{i}(\mathbf{w}))^{2}$.

# Parameter definition for the test
Consider the following parameters:
$$\boldsymbol{\Sigma} = \left[
\begin{array}{ccc}
 1.0000 &  0.0015 & -0.0119\\
 0.0015 &  1.0000 & -0.0308\\
-0.0119 & -0.0308 &  1.0000
\end{array}
\right]$$
$$\mathbf{b} = \left[
\begin{array}{c}
 0.1594 \\
 0.0126 \\
 0.8280
\end{array}
\right]$$
and the corresponding code:
```{r}
Sigma <- rbind(c(1.0000, 0.0015, -0.0119),
               c(0.0015, 1.0000, -0.0308),
               c(-0.0119, -0.0308, 1.0000))
b <- c(0.1594, 0.0126, 0.8280)
```

The optimal solution is known to be
$$\mathbf{w}^\star = \left[
\begin{array}{c}
 0.2799 \\
 0.0877 \\
 0.6324
\end{array}
\right],$$
but the Matlab implementation seems to converge to
$$\mathbf{w} = \left[
\begin{array}{c}
 0.3106 \\
 0.0000 \\
 0.6895
\end{array}
\right].$$
Let's explore this problem with the R package [riskParityPortfolio](https://CRAN.R-project.org/package=riskParityPortfolio).

# Vanilla formulation
This problem is a vanilla formulation because it just contains the risk-parity term subject to the budget constraint and the no-shortselling constraint. Therefore, it can be reformulated as a convex problem and the global optimal solution can be obtained:
```{r, message=FALSE}
library(riskParityPortfolio)

res <- riskParityPortfolio(Sigma, b = b)
res$w
res$risk_contribution/b
```


# Formulation "rc-over-var vs b"
Even though we really have a vanilla formulation, we can still consider a direct nonconvex formulation. Consider the risk expression:
$$R(\mathbf{w}) = \sum_{i=1}^{N}\left(\frac{w_{i}\left(\boldsymbol{\Sigma}\mathbf{w}\right)_i}{\mathbf{w}^T\boldsymbol{\Sigma}\mathbf{w}}-b_i\right)^{2}.$$
The general solver `alabama` is sensitive to the initial point in this formulation (the first case gets stuck in a local minimum):
```{r, message=FALSE}
set.seed(234)
res_ala1 <- riskParityPortfolio(Sigma, b = b, w0 = b, 
                                formulation = "rc-over-var vs b", 
                                method = "alabama")
res_ala1$w
tail(res_ala1$obj_fun, 1)

res_ala2 <- riskParityPortfolio(Sigma, b = b, w0 = c(1, 1, 1)/3, 
                                formulation = "rc-over-var vs b", 
                                method = "alabama")
res_ala2$w
tail(res_ala2$obj_fun, 1)

res_ala3 <- riskParityPortfolio(Sigma, b = b, w0 = (w0 <- runif(3))/sum(w0), 
                                formulation = "rc-over-var vs b", 
                                method = "alabama")
res_ala3$w
tail(res_ala3$obj_fun, 1)
```
```{r}
# plot the portfolios
barplot(rbind(res_ala1$w, res_ala2$w, res_ala3$w),
        main = "Portfolios with different initial points", 
        xlab = "stocks", ylab = "dollars", beside = TRUE, col = heat.colors(3))
# plot the risk contributions
barplot(rbind(res_ala1$risk_contribution/b, res_ala2$risk_contribution/b, res_ala3$risk_contribution/b),
        main = "Risk contribution of the portfolios", 
        xlab = "stocks", ylab = "dollars", beside = TRUE,  col = heat.colors(3))
```

The SCA method gives the same results:
```{r}
set.seed(234)
res_sca1 <- riskParityPortfolio(Sigma, b = b, w0 = b, 
                                formulation = "rc-over-var vs b", 
                                method = "sca")
res_sca1$w
tail(res_sca1$obj_fun, 1)

res_sca2 <- riskParityPortfolio(Sigma, b = b, w0 = c(1, 1, 1)/3, 
                                formulation = "rc-over-var vs b", 
                                method = "sca")
res_sca2$w
tail(res_sca2$obj_fun, 1)

res_sca3 <- riskParityPortfolio(Sigma, b = b, w0 = (w0 <- runif(3))/sum(w0), 
                                formulation = "rc-over-var vs b", 
                                method = "sca")
res_sca3$w
tail(res_sca3$obj_fun, 1)
```
```{r}
# plot the portfolios
barplot(rbind(res_sca1$w, res_sca2$w, res_sca3$w),
        main = "Portfolios with different initial points", 
        xlab = "stocks", ylab = "dollars", beside = TRUE, col = heat.colors(3))
# plot the risk contributions
barplot(rbind(res_sca1$risk_contribution/b, res_sca2$risk_contribution/b, res_sca3$risk_contribution/b),
        main = "Risk contribution of the portfolios", 
        xlab = "stocks", ylab = "dollars", beside = TRUE,  col = heat.colors(3))
```


# Formulation "rc-over-sd vs b-times-sd"
Consider now the risk expression:
 $$R(\mathbf{w}) = \sum_{i=1}^{N}\left(\frac{w_{i}\left(\boldsymbol{\Sigma}\mathbf{w}\right)_i}{\sqrt{\mathbf{w}^T\boldsymbol{\Sigma}\mathbf{w}}}-b_i\sqrt{\mathbf{w}^T\boldsymbol{\Sigma}\mathbf{w}}\right)^{2} = \sum_{i=1}^{N}\left(\frac{r_i}{\sqrt{\mathbf{1}^T\mathbf{r}}}-b_i\sqrt{\mathbf{1}^T\mathbf{r}}\right)^{2}.$$

The general solver `alabama` is again sensitive to the initial point in this formulation (the third case gets stuck in a local minimum):
```{r}
set.seed(234)
res_ala1 <- riskParityPortfolio(Sigma, b = b, w0 = b, 
                                formulation = "rc-over-sd vs b-times-sd", 
                                method = "alabama")
res_ala1$w
tail(res_ala1$obj_fun, 1)

res_ala2 <- riskParityPortfolio(Sigma, b = b, w0 = c(1, 1, 1)/3, 
                                formulation = "rc-over-sd vs b-times-sd", 
                                method = "alabama")
res_ala2$w
tail(res_ala2$obj_fun, 1)

res_ala3 <- riskParityPortfolio(Sigma, b = b, w0 = (w0 <- runif(3))/sum(w0), 
                                formulation = "rc-over-sd vs b-times-sd", 
                                method = "alabama")
res_ala3$w
tail(res_ala3$obj_fun, 1)
```
```{r}
# plot the portfolios
barplot(rbind(res_ala1$w, res_ala2$w, res_ala3$w),
        main = "Portfolios with different initial points", 
        xlab = "stocks", ylab = "dollars", beside = TRUE, col = heat.colors(3))
# plot the risk contributions
barplot(rbind(res_ala1$risk_contribution/b, res_ala2$risk_contribution/b, res_ala3$risk_contribution/b),
        main = "Risk contribution of the portfolios", 
        xlab = "stocks", ylab = "dollars", beside = TRUE,  col = heat.colors(3))
```


The SCA method is also sensitive to the initial point (the first case gets stuck in a local minimum):
```{r}
set.seed(234)
res_sca1 <- riskParityPortfolio(Sigma, b = b, w0 = b, 
                                formulation = "rc-over-sd vs b-times-sd", 
                                method = "sca")
res_sca1$w
tail(res_sca1$obj_fun, 1)

res_sca2 <- riskParityPortfolio(Sigma, b = b, w0 = c(1, 1, 1)/3, 
                                formulation = "rc-over-sd vs b-times-sd", 
                                method = "sca")
res_sca2$w
tail(res_sca2$obj_fun, 1)

res_sca3 <- riskParityPortfolio(Sigma, b = b, w0 = (w0 <- runif(3))/sum(w0), 
                                formulation = "rc-over-sd vs b-times-sd", 
                                method = "sca")
res_sca3$w
tail(res_sca3$obj_fun, 1)
```
```{r}
# plot the portfolios
barplot(rbind(res_sca1$w, res_sca2$w, res_sca3$w),
        main = "Portfolios with different initial points", 
        xlab = "stocks", ylab = "dollars", beside = TRUE, col = heat.colors(3))
# plot the risk contributions
barplot(rbind(res_sca1$risk_contribution/b, res_sca2$risk_contribution/b, res_sca3$risk_contribution/b),
        main = "Risk contribution of the portfolios", 
        xlab = "stocks", ylab = "dollars", beside = TRUE,  col = heat.colors(3))
```

# Conclusion
The vanilla risk-parity portfolio is a convex problem and it can be solved optimally. However, if instead one uses a direct nonconvex formulation (which is required when having different constrains or additional objectives apart from the risk-parity one), there are local minima.


# References {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent



